---
title: "Multiple regression model for Redundancy, Effect traits"
author: "Erick Calderón-Morales"
date: "March 30, 2019"
output: 
  html_document:
    code_folding: hide
    fig_height: 8
    fig_width: 10
    keep_md: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
---
```{r message=FALSE, warning=FALSE}
#Paquetes
library(kableExtra)
library(lattice)
library(tidyverse)
library(PerformanceAnalytics)
library(janitor)
source("C:/tesis_catie/Calderon_CATIE/scripts/functions/HighstatLibV10.R")
library(ggbiplot)
library(GGally)
library(INLA)
```

#Objetivo 
Realizar un modelo linear multiple para los datos de Redundancia funcional
Ref: Capitulo 8 Zuur et al.

```{r}
#Load data
dredun_eff <- read.csv("C:/tesis_catie/Calderon_CATIE/data/resultados_csv/data_indices_eff_redundancy.csv", header=T)
dredun_eff <- clean_names(dredun_eff) %>% 
  select(-c(d))

dparcelas <- read.csv("C:/tesis_catie/Calderon_CATIE/data/clean/data_posicion_parcelas.csv", header = T)
dparcelas <- clean_names(dparcelas)

denv <- read.csv("C:/tesis_catie/Calderon_CATIE/data/clean/data_enviroment_worldclim_clean.csv", header = T, row.names = 1)
denv <- denv %>%
  clean_names() %>% 
    select(-c(crtm_90_x,crtm_90_y,forest_type, slope_per,slope_deg))

#Standarize data
denvi_est <-  scale(denv,center = F, scale = T )
denvi_est <- rownames_to_column(as.data.frame(denvi_est),var= "plot")
```


#Join Data sets

```{r}
dfull  <- left_join(dredun_eff, dparcelas,by=c("plot") ) %>% 
  left_join(.,denvi_est, by=c("plot") )

#Ordenar columnas
dfull_est <- dfull %>% select(plot, forest_type,longitude,latitude,crtm_90_x,crtm_90_y,n,redundancy,u , q, everything()) 
glimpse(dfull_est)
```



#Data exploration
##PCA sin variables de respuesta

```{r}
ncol(denvi_est)
pca_envi <- prcomp(denvi_est[,2:18], scale. = TRUE)
ggbiplot(pca_envi, obs.scale = 3, var.scale = 3,
  groups = dfull_est$forest_type, ellipse = TRUE) +
  scale_color_discrete(name = '') +
  theme(legend.direction = 'horizontal', legend.position = 'top')+
  theme_grey()
```




##Correlaciones Redundancy vrs Environment

```{r fig.width=40, fig.height=40, results='asis', fig.fullwidth=TRUE, fig.dim= 200}
chart.Correlation(dfull_est[,c(8,3,4, 11:27)], histogram=TRUE, pch="+") 
```

##VIFs
```{r}
corvif(dfull_est[,c(3,4,11:27)])
```

##Covariables que se van a eliminar 

```{r}
dredundancy_eff <- dfull_est %>% 
  select(-c(tempsd, mg, ca, p, elev, sand, prec, preccv, tempmin, limo, p_h, u,q, crtm_90_x,crtm_90_y ))
```

+ __tempsd__. Debido a que no tiene relacion con redundancy
+ __mg__. Debido a que no tiene relacion con redundancy
+ __ca__. Debido a que no tiene relacion con redundancy
+ __p__. Debido a que no tiene relacion con redundancy
+ __elev__. Debido a que tiene alto VIF, alta correlacion con Latitude y en el pca tiene la misma direccion que organic matter y sand 
+ __sand__. Debido a que tiene alto VIF  y en el pca tiene la misma direccion que organic matter.
+ __prec__. alta correlacion con predriest y en el pca tiene la misma direccion que prec.
+ __preccv__. Alta correlacion con predriest, en el pca tiene la misma direccion que prec y tiene poca relacion con redundancia.
+ __limo__. Tiene poca relacion con redundancia.
+ __pH__. Tiene poca relacion con redundancia.



###VIFs sin variables eliminadas 
```{r}
ncol(dredundancy_eff)
corvif(dredundancy_eff[,c(3,4,7:12)])
```



##Cleveland plot

```{r}
Mydotplot(dredundancy_eff[, 8:12])
```



#Model formulation


$$redundancia \sim N(\mu_i, \sigma^2)$$

$$E(redundancia)=\mu_i ,var(redundancia)=\sigma^2$$

$$\mu_i= \beta_1*foresttype+\beta_2*long+\beta_3*lat+\beta_4*clay+\beta_5*acidity$$  $$+\beta_6*k+\beta_7*organicmatter+\beta_8*precdriest+\beta_9*temp + \beta_{10}*n $$

```{r}
glimpse(dredundancy_eff)
```

##Model R-INLA
```{r }

mnull <- inla(redundancy ~ longitude + latitude,
           family = "gaussian",
           control.predictor = list(
             compute = TRUE,
             quantiles = c(0.025, 0.975)
           ),
           control.compute = list(dic = TRUE,
                                  waic =TRUE),
           data = dredundancy_eff)
 
m1 <- inla(redundancy ~ forest_type ,
           family = "gaussian",
           control.predictor = list(
             compute = TRUE,
             quantiles = c(0.025, 0.975)
           ),
           control.compute = list(dic = TRUE,
                                  waic =TRUE),
           data = dredundancy_eff)
 


m2 <- inla(redundancy ~ forest_type + clay ,
           family = "gaussian",
           control.predictor = list(
             compute = TRUE,
             quantiles = c(0.025, 0.975)
           ),
           control.compute = list(dic = TRUE,
                                  waic =TRUE),
           data = dredundancy_eff)

mcompleto <- inla(redundancy ~ forest_type + longitude + latitude + clay + acidity +
             k + organic_matter + precdriest + temp,
           family = "gaussian",
           control.predictor = list(
             compute = TRUE,
             quantiles = c(0.025, 0.975)
           ),
           control.compute = list(dic = TRUE,
                                  waic =TRUE),
           data = dredundancy_eff)

mcompleto_interac <- inla(redundancy ~ forest_type * longitude +  forest_type*latitude + forest_type*clay + forest_type*acidity +
             forest_type*k + forest_type*organic_matter + forest_type*precdriest + forest_type*temp,
           family = "gaussian",
           control.predictor = list(
             compute = TRUE,
             quantiles = c(0.025, 0.975)
           ),
           control.compute = list(dic = TRUE,
                                  waic =TRUE),
           data = dredundancy_eff)
```

#Model selection

####DIC
```{r}
mnull$dic$dic
m1$dic$dic
m2$dic$dic
#m3$dic$dic
mcompleto$dic$dic 
mcompleto_interac$dic$dic 
```

####WAIC
```{r}
mnull$waic$waic
m1$waic$waic
m2$waic$waic
#m3$waic$waic
mcompleto$waic$waic
mcompleto_interac$waic$waic

```
##Numerical outputs for the Betas
```{r}
betanull <- mnull$summary.fixed[,c("mean","sd", "0.025quant","0.975quant")]
kable(betanull,digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),full_width = F)
```

```{r}
beta1 <- m1$summary.fixed[,c("mean","sd", "0.025quant","0.975quant")]
kable(beta1,digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),full_width = F)
```


```{r}
beta2 <- m2$summary.fixed[,c("mean","sd", "0.025quant","0.975quant")]
kable(beta2,digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),full_width = F)


```


```{r}
betacompleto_interac <- mcompleto_interac$summary.fixed[,c("mean","sd", "0.025quant","0.975quant")]
kable(betacompleto_interac,digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),full_width = F)


```


##Numerical outputs for the hyperparameters

###Precision($\tau$)


```{r}
kable(m1$summary.hyperpar,digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),full_width = F)
```



###Convertir posterior mean $\tau$ a posterior mean $\sigma$ 

```{r}
tau <- m1$marginals.hyperpar$`Precision for the Gaussian observations`
#m1$marginals.hyperpar[[1]]

MySqrt <- function(x){1/sqrt(x)}
(sigma <- inla.emarginal(MySqrt,tau))
```


##Graphical output for the hyperparameters


###Posterior marginal distribution of $\tau$
```{r}
tau <- m1$marginals.hyperpar$`Precision for the Gaussian observations`
plot(tau[,1], tau[,2], type="l",
     xlab = expression(tau),
     ylab = expression(paste("P(",tau,"|Data)")))

```

###Spline smoothing applied on the posterior marginal distribution of $\tau$
```{r}
tau.distr <- inla.smarginal(tau)
plot(tau.distr$x, tau.distr$y, type="l",
     xlab = expression(tau),
     ylab = expression(paste("P(",tau,"|Data)")))
```


###Posterior marginal distribution of $\sigma$
```{r}
sigma.distr <- inla.tmarginal(fun = MySqrt, 
                              marg = tau)
plot(x = sigma.distr[,1], 
     y = sigma.distr[,2],
     type = "l", 
     xlab = expression(sigma),
     ylab = expression(paste("P(", sigma ," | Data)")))
text(0.86, 8.2, "C", cex = 1.5)
```


#Fitted Model

$$redundancia \sim N(\mu_i, 0.02656408^2)$$

$$E(redundancia)= \mu_i, var(redundancia)=0.02656408^2$$


Solo P. macroloba es importante:

$\mu_i$ = 4.201 + 0.024


#Model validation

##Calculating the residuals in R-INLA

```{r}
Fit <- m1$summary.fitted.values[,"mean"]
E1 <- dredundancy_eff$redundancy - Fit

par(mar = c(5, 5, 2, 2), cex.lab = 1.5)

plot(x = Fit, 
     y = E1, 
     xlab = "Fitted values",
     ylab = "Residuals")

range(dredundancy_eff$redundancy)
par(mar = c(5, 5, 2, 2), cex.lab = 1.5, pty = "s")
plot(x = Fit, 
     xlim = c(0.77, .92),
     ylim = c(0.77, .92),
     y = dredundancy_eff$redundancy, 
     xlab = "Fitted values",
     ylab = "Observed latency scores")

```

```{r}
hist(E1, main = "Normality", breaks=10)
```
```{r}
#Plot residuals versus covariates
plot(x = dredundancy_eff$forest_type, 
     y = E1)
```





#####Importante:
There is the __number of equivalent replicates__. We have a sample size (n) and the model regression parameters (p) . A general recommendation on statistics is to have al least 15 obsevations per parameter (n/p = 15 )

#Visualising the model
























